### 2024-12-05 index.qmd
rearrange

---
title: "100 years of Pies vs. Bars"
format:
  pdf:
    keep-tex: true
    documentclass: article
    papersize: letter
    fontsize: 10pt
    template-partials:
      - title.tex
      - before-bib.tex
    include-in-header:
      - text: |
          \usepackage{sdss2020} % Uses Times Roman font (either newtx or times package)
          \usepackage{url}
          \usepackage{hyperref}
          \usepackage{latexsym}
          \usepackage{amsmath, amsthm, amsfonts}
          \usepackage{algorithm, algorithmic}
          \usepackage[dvipsnames]{xcolor} % colors
          \newcommand{\mt}[1]{{\textcolor{blue}{#1}}}
          \newcommand{\svp}[1]{{\textcolor{RedOrange}{#1}}}
    classoption: [twocolumn]
    mainfont: Times New Roman
    colorlinks: true
author:
  - name: Maksuda Atkar Toma
    affiliation:
      department: Statistics Department
      name: University of Nebraska, Lincoln
    email: mtoma2@huskers.unl.edu
    corresponding: true
  - name: Susan Vanderplas
    affiliation:
      department: Statistics Department
      name: University of Nebraska, Lincoln
    email: svanderplas2@unl.edu
    orcid: 0000-0002-3803-0972
bibliographystyle: acl
bibliography: refs.bib
filters:
   - latex-environment
commands: [mt,svp]
---

# Historical Background

William Playfair introduced the pie and bar charts in the early 1800s[, and by the turn of the next century, strong opinions about the effectiveness of alternative displays of the same data had been publicized in textbooks and manuals.]{.svp}
[For instance, @eellsRelativeMeritsCircles1926 quotes two contemporary resources which discuss pie charts as "not a desirable form of presentation"  [@brintonGraphicMethodsPresenting1914] and more strongly, "an insult to a man's intelligence" [@karsten1923charts].]{.svp}
[@brintonGraphicMethodsPresenting1914 even identifies that the "sector method" is common, but if horizontal bar methods were used more frequently, they would be preferred for both speed and accuracy.]{.svp}
[@eellsRelativeMeritsCircles1926 also quotes an excerpt from @secristIntroductionStatisticalMethods1925, an introductory statistics textbook from 1925, which argues that pie charts require more cognitive effort to use accurately than a comparable stacked bar chart, calling pie diagrams "clumsy and defective" [p 184].]
[These insights mirror today's status quo: pie charts are stubbornly common, even though many design guidelines recommend against their use [@ichartsmarketingdepartment4ProblemsPie2017;@holtzIssuePieChart2018] and a variety of empirical studies identify problems with circular data representations [@hofmannGraphicalTestsPower2012].]{.svp} 
[@wickhamGraphicalCriticismHistorical2013 previously noted the prescience of design guidelines prior to empirical studies validating their recommendations across different types of visualizations. ]{.svp}

[Introspection is an incredibly powerful tool when applied to the design of statistical charts, and can lead to design guidelines that precede empirical tests but match their results. ]{.svp} 
[However, empirical tests are also extremely important: Tufte's data-ink ratio is oft repeated, but empirical studies show that the relationship between minimalism and chart perception is much more complicated [@gillanMinimalismSyntaxGraphs1994a;@tufteVisualDisplayQuantitative2001;@parsonsDataVisualizationPractitioners2020].]{.svp}

Interestingly, the question of pies and bars motivated some of the first empirical evaluations of charts, published in the Journal of the American Statistical Association: @eellsRelativeMeritsCircles1926, @vonhuhnFurtherStudiesGraphic1927, and @croxtonBarChartsCircle1927. 
Experimenters approached the problem head-on, directly asking participants to estimate quantities from charts, and evaluated participants on accuracy, and in some cases, speed. 
While more sophisticated testing methodologies have been developed over the past century [see][@vanderplasTestingStatisticalCharts2020][for a review], studies are still conducted today using direct estimation approaches [@vanderplasFramedReproducingRevisiting2019].

[The initial studies of the question of pies and bars failed to identify a clear "best chart" across all situations, or even to clearly identify in which situations pie charts or bar charts should be preferred.]{.svp} 
[@eellsRelativeMeritsCircles1926 found that pie charts were read as accurately as bar charts and with comparable speed.]{.svp}
[The simplified nature of the experiment attracted critiques from @vonhuhnFurtherStudiesGraphic1927; attached to these criticisms is a short study previously run by Croxton which examined estimates of the ratio between smaller and larger sectors of two-category pie and bar charts; these contributions will be addressed separately even though they are often cited as a single paper, even within the JASA records.]{.svp}
@vonhuhnFurtherStudiesGraphic1927's paper provides several examples which highlight concerns with the Eells study, and argues that while pie charts may be better in certain simple situations, bar charts are much more extensible and facilitate more advanced comparisons, but does not include any experimental studies.
(This is a common issue even today - experimental work is often criticized by those who provide toy examples but no empirical data, as if one person's experience is equivalent to another's experiment.)
Croxton's early experimental data (attached to @vonhuhnFurtherStudiesGraphic1927) had a larger number of participants than @eellsRelativeMeritsCircles1926, but only tested two-category charts with only two proportions: .25 and .4. In addition, Croxton's study elicited estimates from participants as $A/B$ comparisons, where $A + B = 1$, while Eells elicited estimates for $A$ and $B$ separately. 
[It seems that prior to @eellsRelativeMeritsCircles1926, some researchers were empirically testing charts, but these experiments were not necessarily published for broader consumption; when the conversation arose in JASA, Croxton had data from years prior ready to provide additional insight.]{.svp}
This multi-year back-and-forth conversation is interesting in part because it pinpoints an early example of experimental statistical graphics as a discipline of interest to statisticians for both personal investigation and as an audience for the results.

[Subsequently, @croxtonBarChartsCircle1927 published additional data assessing accuracy of estimates from pie and bar charts, varying number of categories and orientation/alignment of sections.]{.svp}

[@croxtonBarChartsCircle1927 found that pie charts may match or surpass them in certain situations.]{.svp}
[Much later, @clevelandGraphicalPerceptionTheory1984 conducted experiments on simple graphical elements, stating that aligned length judgments are more accurate than judgments based on area or angle, with the implication that bar charts are more accurately perceived than pie charts.]{.svp}
[Their hierarchy of feature comprehension is often assumed to be experimentally derived, but the experiments presented in the papers [@clevelandGraphicalPerceptionTheory1984; @clevelandGraphicalPerceptionGraphical1985; @clevelandExperimentGraphicalPerception1986] only covered a few elements of this hierarchy.]{.svp}

[Here, we re-examine the historical literature surrounding the use of pie and bar charts, examining recommendations about the use of pie and bar charts, tracing these guidelines back to empirical studies, and evaluating whether the recommendations are well-supported by the studies.]{.svp}
[We begin with an initial set of four studies from the late 1920s and early 1930s: @eellsRelativeMeritsCircles1926, @croxtonBarChartsCircle1927, @vonhuhnFurtherStudiesGraphic1927, and @croxtonGraphicComparisonsBars1932.]{.svp}
[We trace these studies forward in time, examining the claims made about the studies as well as the recommendations made in citing papers.]{.svp}
[In addition, we lay out the different experimental designs used in each experimental study, with the goal of assessing how the design of each study might contribute to the ultimate results.]{.svp}