### 2025-06-17 presentation.qmd
Added readme and presentation files

---
title: "100 years of Pies vs. Bars"
author: " Maksuda Aktar Toma"
institute: "University of Nebraska Lincoln"
date: today
format:
  beamer:
    theme: metropolis
    colortheme: beaver
    title-slide-attributes: "align=center"
    pdf-engine: xelatex
    mainfont: Times New Roman
---

# Content

- Introduction
- Motivation
- The Dawn of Empirical Graphics
- Modern Experimental Graphics
- Historical Experimental design
- Result
- Conclusion

# Introduction: Pie vs. Bar — A Century-Long Debate

- Pie and bar charts date back to Playfair (early 1800s).
- Pie charts often criticized as less accurate, yet remain widely used.
- This review explores:
  - Historical experiments and design
  - Task performance comparisons
  - Practical usefulness of each chart type

# Motivation: Why Revisit Pie vs. Bar?

- Effective charts are essential for accurate communication of quantitative data.
- Bar charts consistently outperform pie charts in accuracy, speed, and clarity (based on early research).
- Yet, pie charts remain popular in media, business, and education.
  - This raises key questions:
  - Why do pie charts persist?
  - How do habits, aesthetics, or tradition influence chart selection?

![Pie vs. Bar](000023.png){width=80% height=160px}

# The Dawn of Empirical Graphics

- Early JASA studies (1926–1930s) by Eells, von Huhn, and Croxton tested pie vs. bar charts for accuracy and speed.
- Tasks varied: Eells used part estimation; Croxton used ratio comparisons like $A/B$, $A+B=1$
- Findings: No clear winner—results depended on task design; von Huhn and Croxton critiqued and refined earlier work.

![*Examples of comparison tasks used in graphical perception studies*](pie-slices.png)

# Modern Experimental Graphics

![*Visual encodings ordered by perceptual accuracy, as proposed by Cleveland and McGill*](cleveland-mcgill.jpg){width=80% height=200px}

\scriptsize
- The first six activities are listed according to Cleveland and McGill's accuracy tests. Assumed to be far less accurate than the others.

# Modern Experimental Graphics (Cont.)

- Spence & Lewandowsky (1991): Compared A vs. B+C using bar, pie, and table; measured accuracy and speed.
- Peterson (1954): Evaluated 8 chart types for part-whole accuracy; found format matters.
- Simkin & Hastie (1987): Bars better for comparisons; pies better for proportions.
- Skau & Kosara (2016): Pie distortions reduce accuracy; arc cues work best.
- Hill (2025): Bars best for ranking; pies/donuts strong in simple proportion tasks.

# Experimental Description

![*Summary of Experimental Studies on Chart Perception*](picture3.png){height=280px}

# Accuracy of Charts

![*Accuracy of Bar Chart By Task Type*](000013.png){width=80% height=200px}

\scriptsize
- Filled dots indicate studies that reported standard errors; hollow dots did not
- Hill (2024) showed the lowest estimation error, indicating the highest accuracy among all studies.
- Studies with Close bar alignment performed better than those with Far bar placement

# Accuracy of Charts (Cont.)

- Heer & Bostock (2010) showed the lowest error, indicating better accuracy with segmented bars.
- Older studies had higher error rates and often lacked standard error reporting (hollow points).

:::{.columns}
::: {.column width="50%"}

![*Accuracy of Bar Chart By Task Type*](000013.png)
:::
::: {.column width="50%"}

![*Accuracy of Bar Chart By Task Type*](00009.png)
:::
:::

# Accuracy of Charts (Cont.)

- Pie charts perform well for part-to-whole tasks, with low error (1–3%).
- Segment comparisons are harder, showing higher errors (4–6%).
- Recent studies show improved accuracy with better design.

:::{.columns}
::: {.column width="50%"}

![*Accuracy of Pie Chart By Task Type*](00008.png)
:::
::: {.column width="50%"}
![*Accuracy of Pie Chart By Task Type*](00007.png)
:::
:::

---

# Result

- Bar charts excel in comparison and ranking tasks — faster, more accurate, and user-preferred.
- Pie charts perform well in part-to-whole and grouped estimation tasks, aligning with intuitive reasoning.
- Task matters: Bar > Pie for $A$ vs $B$; Pie > Bar for $A / (A+B)$; mixed results for $A$ vs $B+C$, $A+B$ vs $C+D$

# Conclusion

:::{.callout-tip title="Conclusion"}
There is no universal winner.
The most effective chart type depends on the specific judgment or task.
:::

- Both chart types remain widely used across science, business, and social sciences.
- The effectiveness of a chart depends on:
  - the task (comparison, estimation, ranking),
  - perceptual features (angle, length, position),
  - and visual design elements ( adjacency, clarity).
- Pie charts are generally better for conveying part-to-whole relationships.
- Bar charts are superior for comparing magnitudes and identifying rankings.

# References

\scriptsize

Funkhouser, H. G. (1937). Historical development of the graphical representation of statistical data. Osiris, 3, 269–404. https://doi.org/10.1080/03697827.1937.10559475

Croxton, F. E., & Stein, H. (1932). Graphic comparisons by bars, squares, circles, and cubes. Journal of the American Statistical Association, 27(177), 54–60.

von Huhn, R. (1927). Further studies in the graphic use of circles and bars: A discussion of the Eells experiment. Journal of the American Statistical Association, 22(157), 31–39. https://doi.org/10.2307/2277346

Croxton, F. E., & Stryker, R. E. (1927). Bar charts versus circle diagrams. Journal of the American Statistical Association, 22(160), 473–482. https://doi.org/10.2307/2276829

Eells, W. C. (1926). The relative merits of circles and bars for representing component parts. Journal of the American Statistical Association, 21(154), 119–132. https://doi.org/10.2307/2277140

Cleveland, W. S., & McGill, R. (1984). Graphical perception: Theory, experimentation, and application to the development of graphical methods. Journal of the American Statistical Association, 79(387), 531–554. https://doi.org/10.1080/01621459.1984.10478080

Cleveland, W. S., & McGill, R. (1985). Graphical perception and graphical methods for analyzing scientific data. Science, 229(4716), 828–833. https://doi.org/10.1126/science.229.4716.828

# References

\scriptsize

Cleveland, W. S., & McGill, R. (1986). An experiment in graphical perception. International Journal of Man-Machine Studies, 25(5), 491–500. https://doi.org/10.1016/S0020-7373(86)80019-0

Simkin, D., & Hastie, R. (1987). An information-processing analysis of graphs and tables. Cognitive Psychology, 19(4), 530–544. https://doi.org/10.2307/2289447

Holtz, Y. (2018, August). The issue with pie chart. Data To Viz. https://www.data-to-viz.com/caveat/www.data-to-viz.com/caveat/pie.html

iCharts Marketing Department. (2017, November). 4 problems with pie charts (and what to do instead). iCharts. https://web.archive.org/web/20180223231734/https://icharts.net/blog/data-visualization-guides/4-problems-pie-charts-instead

Skau, D., & Kosara, R. (2016). Arcs, angles, or areas: Individual data encodings in pie charts. Computer Graphics Forum, 35(3), 121–130. https://doi.org/10.1111/cgf.12888

Heer, J., & Bostock, M. (2010). Crowdsourcing graphical perception: Using Mechanical Turk to assess visualization design. In Proceedings of the ACM CHI Conference on Human Factors in Computing Systems (pp.203–212). https://doi.org/10.1145/1753326.1753357

Hill, N. D. (2025). Bar, pie or donut? Exploring chart effectiveness for part-to-whole communication. Information Visualization, 24(1), 3–20. https://doi.org/10.1177/14738716241259432

---

# Thank You

::: {style="text-align: center;"}

\begin{center}
\LARGE \textbf{Questions?}
\end{center}

:::